### 2011-10-03 doc/manuscript.tex
The posterior predictive checking procedure is being done with s17 simulation.

phosphotransferase system (PTS), ATP-binding cassette (ABC), Aminoacyl
tRNA synthetase (aaRS), ancestral recombination graph (ARG), group
A streptococci (GAS), group C streptococci (GCS), group G streptococci
(GGS), \textit{Streptococcus pyogenes} (SPY), \textit{S.\ dysgalactiae}
subspecies \textit{dysgalactiae} (SDD), \textit{S.\ dysgalactiae} subspecies
\textit{equisimilis} (SDE).

@@

\section*{Abstract}
Homologous recombination and horizontal gene transfer are investigated with
complete genomes of closely related bacterial species of \textit{Streptococcus
pyogenes} (SPY) and \textit{S.\ dysgalactiae} subspecies \textit{equisimilis}
(SDE).  The former is a pervasive human-specific pathogen, and the latter an
opportunistic human pathogen.  Parts of the genomic configuration, including
human pathogenicity, of SDE could have originated in those of SPY.  Our
genome-wide survey of gene transfer indicated that gene transfer via homologous
recombination occurred relatively from SPY to SDE.  Although we did not find
enrichment of virulence genes in the homologous recombining gene transfer, we
found evidence that they transferred more often via non-recombining fashion than
non-virulence genes.  A genome-wide resource of recombining and non-recombining
gene transfer for the study of the genomes is provided so that research
community in pathogenic bacteria can take advantage of it.

@@

The research topic on gene exchanges between SPY and SDE has been discussed in
various ways. \citet{Sachse2002} showed gene transfer from SDE to SPY for the
downstream of the genomic region of streptococcal pyrogenic exotoxins
\textit{speG}. \citet{Davies2007} investigated phage 3396 from SDE as a vehicle
of gene transfer from SDE to SPY. \citet{Davies2009} demonstrated that an
integrative conjugative element of SDE allowed gene transfer from SDE to other
streptococci.  \citet{Ahmad2009} did not conclusively deduce the net
directionality of housekeeping gene transfer between SPY and SDE.  Notably,
\citet{Kalia2001} attempted to show that the direction of gene transfer from SPY
to SDE was more dominant than the other direction.  Because these researches on
the controversial issue, directionality of gene transfer, have been limited to
relatively small regions in the genomes, it remains to be elusive to determine
the genome-wide degrees to which genomic segments flow between the two species,
and directions of thereof.

As sequencing costs plummet \citep{Mardis2011}, bacterial genome sequencing
produces individual genomes of high quality even within a bacterial species
\citep{Tettelin2009a}.  Studies on horizontal gene transfer in prokaryotes have
used genes in very large sets of species \citep{Koonin2001} from a great deal
of genomes available.  Although this type of study gave us insights into complex
evolutionary histories of prokaryotes, with multiple individual genomes per
species we now can go beyond the broad study on gene transfer towards
fine-scaled history of gene transfer between very closely related prokaryotes.
For instance, \citet{Liu2009} studied gene transfer events in
\textit{Lactobacillus bulgaricus} and \textit{Streptococcus thermophilus} using
the compositional approach developed by \citet{Karlin2001}. \citet{Luo2011}
investigated gene transfer events in \textit{Eschericha Coli} using the embedded
quartet decomposition analysis developed by \citet{Zhaxybayeva2006}.
\citet{Hamady2006} combined phylogenetic and compositional approaches to
detecting horizontal gene transfer.  \citet{Caro-Quintero2011} studied a number
of \textit{Shewanella baltica} genomes to identify the genetic elements that
enabled the species to adapt to redox gradients.

Bacteria evolves new biological features through horizontally transferred genes
from other lineages and vertically transmitted mutations on genes within
the same lineage.  Mutations on a gene modifiy relatively small portions of the
gene, including substitutions, insertions, and deletions of deoxyribonucleic
acids. Horizontal gene transfers can reshape the organization of a bacterial
genome, which could revamp biological functions of the bacterial species.
Recipient bacteria can substitute transferred genes for homologous parts of
their genomes. They can also integrate transferred genes into their genomes
without replacement. These two processes are here discriminated by calling the
former \textit{recombining gene transfer}, and the latter \textit{horizontal
gene transfer} \citep{Ochman2001,Lawrence2009}.  Despite the two distinct
evolutionary processes in bacteria, reseraches on statistical analyses of
complete genome sequences in efforts of better understanding bacterial evolution
have paid little attention to a combined approach to illuminating disparity of
the two gene transfer processes.

Homologous recombining gene transfer could be studied more readily using the
core genomes from multiple individual bacterial genomes while horizontal gene
transfer might be more easily detected using the pan genomes. Recently, a method
of inferring homologous recombination with whole bacterial genomes was developed
\citep{Didelot2010}. The model of homologous recombination allowed them to study
flows of genetic segment between bacterial individuals of \textit{Bacillus
cereus} species in the core genomes.  More complex evolutionary processes acting
on bacterial genomes include genome rearragement, translocation, inversion,
duplication, gain, loss, etc.  Because of the complexity, these processes do not
readily lend themselves to statistically more sound approaches such as maximum
likelihood. Instead, a parsimony-based approach to inferring gene gain, loss,
and transfer events was developed by  \citep{Doyon2011}.  We noted that the
first approach could be used to infer transfer of genomic segments due to
homologous recombination, and the second could do similarly due to horizontal
gene transfer. We expect that model-based approaches using whole bacterial
genomes will be developed in the near future.

The complete genomes of the two spcies of SPY and SDE presented a unique
opportunity to study gene transfer over relatively recent time scales.  We used
both of the model-based approach for inferring homologous recombination and the
parsimony-based one for detecting gene gain, loss, and horizontal transfer. We
also compared both of the results in recombining gene transfer events to show
that the parimony-based method inferred gene transfer events in similar
biological functional categories to those inferred with the model-based
approach. This is the first attempt to analyaze SPY and SDE in genome-wide scale
for detecting gene transfer events not only in the core genome but also in the
pan genome, which make the analysis complete in terms of the amount of used
genome data. Our analysis of core genomes also showed for the first time that
the net directionality of, at least, recombining gene transfer was more from SPY
to SDE.

@@

\subsection{The five or six genomes}

\subsection{The genome alignment of the five genomes}

\subsection{The species tree inferred using the 274 alignment blocks}

\subsection{The homologous recombination events}
An alignment of the five genomes showed two distinct genome structures of SPY
and SDE (Figure S\ref{fig:mauve}). The SDE genomes were larger in length than
those of SPY. Although it was unclear that SDD was different from either of SDE1
and SDE2 from the graphical view of the genome alignment, the spcies tree from
the genome alignment showed that SDD branch was attached to the clade branch of
SDE (Figure S\ref{fig:tree5}). The distribution of the alignment block lengths
was skewed to the right with mean of 4188 base pairs long (Figure
S\ref{fig:blocksize}), which implied that the model assumption of geometric
length distribution might be plausible in \texttt{ClonalOrigin} method.  The
relative strength of recombination compared to mutations indicated that
reconmination-driven substitutions affected more sites than did mutation-driven
substitutions (Table S\ref{tab:clonalframe}).

\subsubsection{The three population parameters}
Along the genome of SDE1 population mutation rate did not appear to change much
(Figure \ref{fig:scatter3}).  We found a few spots of genomic regions where the
mutation rates were larger than twice the global estimate of mutation rate.
\texttt{WHAT ARE THESE BLOCKS? WHAT GENES ARE LOCATED IN THE BLOCK?}
The recombination rate appeared to change more dynamically. Espeically, the
genomic region of SDE1 near 20 kilobase pairs from the origin of replication
contains alignment blocks with bell-shaped recombination rates.
\texttt{USE UCSC GENOME BROWSER TO FIND WHAT GENES BELONG TO THOSE BLOCKS.}
The recombinant tract length changes in a way that we could not note any
particular positions along the genome.
There appeared to be negative correlation between recombinant tract length and
recombination rate.
\texttt{PLOT PAIRWISE SCATTER PLOTS OF THE THREE PARAMETERS.}
In the first stage of \texttt{ClonalOrigin} MCMC a few alignment blocks were not
successfully finished even after a month of computation.  Because those were not
expected to affect the global estimates of population parameters significantly,
we used only finished results of the first stage of \texttt{ClonalOrigin} to
estimate mutation rate, recombination rate, and recombinant tract length.  Two
independent runs of the first stage MCMC resulted in similar estimates of the
three population parameters (Table S\ref{tab:three}).

\subsection{Tree topology changes along the genome}
Most of local tree topologies, or 66.7\%, along the alignment blocks were the
same as the species tree, which allowed us to be more confident of the species
tree inferred using the alignment blocks (Table
S\ref{tab:Gene-tree-topologies}).  The second most probable topology was one
where SDD was the outgroup, and SPY and SDE were sister clades with probability
of 9.3\%.  Therefore, SPY and SDE appeared to share about tenth of the core
genomes that were more closely related with each other than each to SDD's
genome. The other topologies had less than 4\% of proportions.

\subsection{Evidence for unbalaned gene flow from SPY to SDE}
The heat map in Figure \ref{fig:Heatmap-of-recombination} shows the log base 2
ratios of these numbers and their corresponding expected values given the
species tree and the global parameter settings.  Because the recombinant edge
from any species tree branch to the root branch is not possible under
\texttt{ClonalOrigin} model (see Figure \ref{fig:clonalorigin}), the column for
the root as recipient was colored green. Likewise, all of the cells where a
recipient branch was older than the donor, and two branches did not share any
contemporary time were colored green. The diagonal cells were colored gray
because the observed number of recombinant edges and the expected one were
similar. Because recombinant edges starting from and ending in the same species
tree branch would not change the species tree, which would not change the
likelihood, the cells should be colored gray.  Because SDD is strictly
veterinary hosts, we did not expect that SDD received much gene transfer from
either SPY1 or SPY2.  The log base 2 ratio from SPY1 to SDD was -2.20, and that
from SPY2 to SDD was -1.97. On the contrary, the recombing gene transfer from
SDE1 to SDD was 0.756, and that from SDE2 to SDD was 0.698. The relatively
higher recombining gene transfer from two SDE individuals to SDD compared with
that from two SPY individuals to SDD reflected that the SDD was more closely
related to SDE than to SPY.  The gene transfer from SDD to either SDE1 or SDE2
was less frequent than the case of reverse direction.

The recombining gene transfer log base 2 ratios between SDE1 and SDE2 were about
twice as large as that from SDE1 and SDE2 to SPY1 and SPY2. Because the
recombining gene transfer within a clade would be larger than that between
clades, this was expected. Surprisingly, the recombining gene transfer log base
2 ratios between SPY1 and SPY2 was comparable with that from SPY1 and SPY2 to
SDE1 and SDE2.  The recombining gene transfer from a clade SPY1 or SPY2 to a
different clade SDE1 or SDE2 was not different from that within the clade, or
SPY1 and SPY2.  Consistent with this result, the gene transfer log base 2 ratio
from SPY1 and SPY2 to SDE branch was larger than any type of gene transfer: 1.88
and 1.94, respectively. This clearly showed that the recombining gene transfer
was biased toward the net direction from SPY to SDE.

Furthermore, we found a total of 58 genes 
for which a recombinant edge 
of certain types were sampled with probability at least 0.9
by focusing on specific types of recombinant edges.
(Table S\ref{tab:genes-transfer}).  Many of the genes in the list were
transferred from SPY to SDE rather than the reverse direction overall.

% \texttt{IG: ADD FIGURE WITH NON-NORMALIZED VALUES? SC: Why?}

% No evidence for unbalanced non-recombining gene transfer between SPY and SDE
Figure S\ref{fig:hgt-heatmap} shows the number of inferred
non-recombining gene transfer events between the different branches of
the species tree.
SDE1 and SDE2 shared the most events (26 and 34) compared to other sample pairs. 
SDD appeared to acquire many genes from the ancestral SDE branch. Transfer 
events between the SPY and SDE clades were inferred in smaller numbers (<6) 
compared to those within each clade (23--34). Among the clade-crossing 
transfers, we did not see any significant enrichment for a specific direction, 
as opposed to the result for recombining transfers from the model-based 
analysis (Figure S\ref{fig:mowgli-recomb-heatmap}).

\subsection{The virulence genes}
\texttt{Find cases of any virulence genes that were transferred via recombining and
horizontally.} 
\texttt{I need to see how the virulence genes list was created.}
\texttt{I want to know the virulence gene categories. What kind of virulence
genes are more under gene transfer?}
\texttt{Make a list of virulence genes and their categories with more
interpretations.}

A list of virulence genes that might be transferred via recombining gene
transfer.
block 76: SDEG\_0641
block 97: SDEG\_0673
block 101: SDEG\_0739
block 151: SDEG\_1447
block 244: SDEG\_2040
block 242: SDD27957\_00955
block 240: SDD27957\_00985
block 67: SDD27957\_03085
block 133: SDD27957\_04505
block 167: SDD27957\_06790
block 170: SDD27957\_07725
block 186: MGAS10750\_Spy1444
block 255: MGAS10750\_Spy1922

Blocks with virulence genes.

Check virulence genes (they were combined) at 
SDD27957\_02890, SDD27957\_02895
MGAS10750\_Spy0455, MGAS10750\_Spy0456 (block 75)

\subsection{Horizontal gene transfer is associated with virulence genes}

@@

\subsection{Parsimony-based approach for gene gain, loss, and duplication}
% Gene gain, loss, duplication, and non-recombining horizontal gene transfer
Among the 2314 gene families
the largest family consisted of 59 genes, while
nearly half of gene families (1066) had exactly one gene per species
, or six genes total (Figure S\ref{fig:famsize}).
Figure \ref{fig:Gene-duplication-loss} shows the
inferred gene counts for each extant and ancestral species. Overall,
we inferred 304 duplications, 944 recombining gene transfers, and 294
non-recombining gene transfers, comprising 23.7\% of total transfer
events.  We found that duplications made up only a small fraction of
the events within the clade (10.2\%), while horizontal transfers and
losses were significantly more frequent (41.7\% and 48.0\%,
respectively). We found that duplications were most often
species-specific. Many gene losses likely represented horizontal
transfers that replaced the native gene sequence.  The species SDD and
SDE were the most expanded, both in terms of duplications and in-bound
transfers.

@@

\subsection{Complete bacterial genomes}
Two genomes of \textit{S.\ pyogenes} (SPY), and two of \textit{S.\ dysgalactiae}
ssp.\textit{\ equisimilis} (SDE) were downloaded from NCBI with the following
accessions: NC\_004070 {Beres2002}, NC\_008024 \citep{Beres2006}, NC\_012891
\citep{Shimomura2011}, and CP002215 \citep{Suzuki2011}. We refer to these
samples as SPY1, SPY2, SDE1, and SDE2, respectively. One genome of 
\textit{S.\ dysgalactiae} ssp.\textit{\ dysgalactiae} was provided by 
\citet{Suzuki2011}
with accession identifiers of SddyATCC27957, referred to here as SDD. We also
downloaded the genome (NC\_012471) of \textit{S.\ equi} ssp.\textit{\ equi}
strain 4047 referred to as SEE.

@@

Two stages of MCMC with \texttt{ClonalOrigin}, subversion r19 that had been
downloaded from \url{http://clonalorigin.googlecode.com/svn/trunk}, were
employed independently for each of the 274 alignment blocks as described in
\citet{Didelot2010}.   In the first MCMC we obtained estimates of mutation rate,
recombination rate, and recombinant tract length.  Two independent chains were
executed with $10^6$ burn-in generations and $10^7$ sampling generations,
sub-sampled every $10^5$-th generations. One of the two chains was used to
obtain the block-wise mean estimates, and the other chain was used to validate
convergence of the first stage of MCMC.  The estimators of the three population
parameters were medians weighted over block lengths, which down-weighed less
reliable estimates obtained from shorter alignment blocks.  The second stage of
MCMC with \texttt{ClonalOrigin} was executed with the three population
parameters fixed to the estimated values in the first stage.  Two independent
chains of the second stage of \texttt{ClonalOrigin}'s MCMC were also executed
with a total generation being $1.1\times10^8$. The first $10^7$ generations of
each chain were discarded as burn-in. The remaining $10^8$ generations were
sub-sampled every $10^5$-th generation.

@@

% Figure 5.
%=============================================================================
% FIGURE - Transfer heatmap
\begin{figure}
\begin{center}
\includegraphics[width=5in]{figures/trans-hgt-heatmap}
\end{center}
\vspace{-1.2in}
\caption{Distribution of horizontal gene transfers.}
\label{fig:hgt-heatmap}
\end{figure}
%=============================================================================

@@

% Table S3. Simulation study for the three population parameters.
\begin{table}
\caption{\label{tab:sim-three-population-parameters}
Simulation study of the three population parameters.}
\noindent \begin{centering}
\begin{tabular}{ccc}
 & True & Estimates (Standard Deviation)\tabularnewline
\hline
Mutation rate per site & 0.081 & 0.081 (0.0011)\tabularnewline
Recombination rate per site & 0.012 & 0.011 (0.0003)\tabularnewline
Recombinant tract length & 744 bp & 919 bp (28)\tabularnewline
\end{tabular}
\par\end{centering}
\end{table}
\clearpage{}

>>>>

group A streptococci (GAS), group C streptococci (GCS), group G streptococci
(GGS), \textit{Streptococcus pyogenes} (SPY), \textit{S.\ dysgalactiae}
subspecies \textit{dysgalactiae} (SDD), \textit{S.\ dysgalactiae} subspecies
\textit{equisimilis} (SDE)

@@

\section*{Abstract}
Homologous recombination and horizontal gene transfer are investigated with
complete genomes of closely related bacterial species of \textit{Streptococcus
pyogenes} (SPY) and \textit{S.\ dysgalactiae} subspecies \textit{equisimilis}
(SDE).  The former is a pervasive human-specific pathogen, and the latter an
opportunistic human pathogen.  We aims at which parts of the genomic
organization, including human pathogenic virulence factors, of SDE could have
originated in those of SPY, or vice versa.  Our genome-wide survey of gene
transfer indicated that gene transfer via homologous recombination occurred
relatively from SPY to SDE.  Although we did not find enrichment of virulence
genes in the homologous recombining gene transfer, we found evidence that they
transferred more often via non-recombining fashion without replacing parts of
destination genomes than did non-virulence genes.  A genome-wide resource of
recombining and non-recombining gene transfer for the study of the genomes is
provided so that research community in pathogenic bacteria could take advantage
of it.

@@

Therefore, the research topic on gene exchanges between SPY and SDE has been
discussed in various ways. \citet{Sachse2002} showed gene transfer from SDE to
SPY for the downstream of the genomic region of streptococcal pyrogenic
exotoxins \textit{speG}. \citet{Davies2007} investigated phage 3396 from SDE as
a vehicle of gene transfer from SDE to SPY. \citet{Davies2009} demonstrated that
an integrative conjugative element of SDE allowed gene transfer from SDE to
other streptococci.  \citet{Ahmad2009} did not conclusively deduce the net
directionality of housekeeping gene transfer between SPY and SDE.  Notably,
\citet{Kalia2001} attempted to show that the direction of gene transfer from SPY
to SDE was more dominant than the other direction.  Because these researches on
the controversial issue, directionality of gene transfer between SPY and SDE,
have been limited to relatively small regions in the genomes primarily on
representative loci from the Multi Locus Sequence Typing database, it remains to
be elusive to determine the genome-wide degrees to which genomic segments flow
between the two species, and net directions of thereof.

As sequencing costs plummet \citep{Mardis2011}, bacterial genome sequencing
produces individual genomes of high quality even within a bacterial species
\citep{Tettelin2009a}.  Studies on horizontal gene transfer in prokaryotes have
used genes in very large sets of species \citep{Koonin2001} from a great deal
of genomes available.  Although this type of study gave us insights into complex
evolutionary histories of prokaryotes, with multiple individual genomes per
species we now can go beyond the broad study on gene transfer towards
fine-scaled history of gene transfer between very closely related prokaryotes.
For instance, \citet{Liu2009} studied gene transfer events in
\textit{Lactobacillus bulgaricus} and \textit{Streptococcus thermophilus} using
the compositional approach developed by \citet{Karlin2001}. \citet{Luo2011}
investigated gene transfer events in \textit{Eschericha Coli} using the embedded
quartet decomposition analysis developed by \citet{Zhaxybayeva2006}.
\citet{Hamady2006} combined phylogenetic and compositional approaches to
detecting horizontal gene transfer.  \citet{Caro-Quintero2011} studied a number
of \textit{Shewanella baltica} genomes to identify the genetic elements that
enabled the species to adapt to redox gradients.  Although several complete
genomes from SPY have been available, \citet{Suzuki2011} recently added one more
complete genome of SDE to the previous set of a unique SDE genome. This set of
SPY and SDE genomes prompted us to study genome-wide gene flows between the two
species that afflicted human beings around the globe.

Bacteria evolves new biological features through horizontally transferred genes
from other lineages and vertically transmitted mutations on genes within
the same lineage.  Mutations on a gene modifiy relatively small portions of the
gene through substitutions, insertions, and deletions of deoxyribonucleic
acids. Horizontal gene transfers can reshape the organization of a bacterial
genome, which could revamp biological functions of the bacterial species.
Recipient bacteria of transferred genes can substitute transferred genes for homologous parts of
their genomes, or integrate transferred genes into their genomes
without replacement. These two processes are here in this paper discriminated by calling the
former \textit{recombining gene transfer}, and the latter \textit{horizontal
gene transfer} \citep{Ochman2001,Lawrence2009}.  Despite the two distinct
evolutionary processes in bacteria, reseraches on statistical analyses of
complete genome sequences in efforts of better understanding bacterial evolution
have paid less attention to a combined approach to illuminating disparity of
the two gene transfer processes. We attempted here to use and compare two
approaches for distinguishing recombining gene transfer process from horizontal, or
non-recombining, gene transfer one, and more systematically tackled their roles
in association with virulence genes.

Homologous recombining gene transfer could be studied more readily using the
core genomes from multiple individual bacterial genomes while horizontal gene
transfer might be more easily detected using the pan genomes. Recently, a method
of inferring homologous recombination with whole bacterial genomes was developed
\citep{Didelot2010}. The model of homologous recombination allowed them to study
flows of genetic segment between bacterial individuals of \textit{Bacillus
cereus} species in the core genomes.  More complex evolutionary processes acting
on bacterial genomes include genome rearragement, translocation, inversion,
duplication, gain, loss, etc.  Because of the complexity, these processes do not
readily lend themselves to statistically more sound model-based approaches such
as maximum likelihood. Instead, a parsimony-based approach to inferring gene
gain, loss, and transfer events was developed by  \citep{Doyon2011}.  We noted
that the first approach could be used to infer transfer of genomic segments due
to homologous recombination, and the second could do similarly attributed to
horizontal gene transfer. Although this heterogeneously combined approach could
be unsatisfactory, we hoped to take the unique opportunity of new genomes to
study gene transfer in genome-wide scale among SPY and SDE individual bacteria.

In short, the complete genomes of the two spcies of SPY and SDE presented a unique
opportunity to study gene transfer over relatively recent time scales.  We used
both of the model-based approach for inferring homologous recombination and the
parsimony-based one for detecting gene gain, loss, and horizontal transfer. We
also compared both of the results in recombining gene transfer events to show
that the parimony-based method inferred gene transfer events in similar
biological functional categories to those inferred with the model-based
approach. This is the first attempt as far as we know to analyaze SPY and SDE in
genome-wide scale for detecting gene transfer events not only in the core genome
but also in the pan genome, which make the analysis complete in terms of the
amount of used genome data. Our analysis of core genomes also showed for the
first time that the net directionality of, at least, recombining gene transfer
from SPY to SDE was stronger than the reverse direction.

@@

\subsection{Recombination-driven substitutions were stronger than mutation-driven ones}

The genome alignment of the five genomes (Table S\ref{tab:genome}) showed two distinct
genome structures of SPY and SDE (Figure S\ref{fig:mauve}). The SDE genomes were
larger in length than those of SPY. Although it was unclear that SDD was
different from either of SDE1 and SDE2 from the graphical view of the genome
alignment, the spcies tree inferred from the genome alignment showed that SDD
branch was attached to the clade branch of SDE1 and SDE2 (Figure
S\ref{fig:tree5}). The distribution of the alignment block lengths was skewed to
the right with mean of 4188 base pairs long (Figure S\ref{fig:blocksize}), which
implied that the model assumption of geometric length distribution might be
plausible in \texttt{ClonalOrigin} method (see section \textbf{Materials and
Method} and \citealt{Didelot2010}).  The relative strength of recombination
compared to mutations indicated that reconmination-driven substitutions affected
more sites than did mutation-driven substitutions (Table
S\ref{tab:clonalframe}). In the \textit{Streptococcus} bacteria homologous
recombination appeared to be a stronger evolutionary force than point mutations 
process.

\subsubsection{Recombination hotspots were observed in genomes}

Along the genome of SDE1 population mutation rate did not appear to change much
(Figure S\ref{fig:scatter3}).  We found a few spots of genomic regions where the
mutation rates were larger than twice the global estimate of mutation rate.
% \texttt{WHAT ARE THESE BLOCKS? WHAT GENES ARE LOCATED IN THE BLOCK?}
The recombination rate appeared to change more dynamically. Espeically, the
genomic region of SDE1 near 20 kilobase pairs from the origin of replication
contains alignment blocks with bell-shaped recombination rates.
% \texttt{USE UCSC GENOME BROWSER TO FIND WHAT GENES BELONG TO THOSE BLOCKS.}
The recombinant tract length changes in a way that we could not note any
particular positions along the genome.
There appeared to be negative correlation between recombinant tract length and
recombination rate.
% \texttt{PLOT PAIRWISE SCATTER PLOTS OF THE THREE PARAMETERS.}

\subsection{SPY and SDE were more closely related in about tenth of the genome}

Most of local tree topologies (see section \textbf{Materials and Methods}), 
or 66.7\%, along the alignment blocks were the
same as the species tree, which allowed us to be more confident of the species
tree inferred using the alignment blocks (Table
S\ref{tab:Gene-tree-topologies}).  The second most probable topology (9.3\%)  was one
where SDD was the outgroup, and SPY and SDE were sister clades 
SPY and SDE appeared to share about tenth of the core
genomes that they were more closely related with each other than each to SDD's
genome. The other topologies had less than 4\% of proportions.

\subsection{Unbalaned gene flow from SPY to SDE in genome-wide scale}

The heat map in Figure \ref{fig:Heatmap-of-recombination} shows the log base 2
ratios of the \textit{a posteriori} number of recombinant edges for a pair of
donor and recipient species tree branches and their corresponding expected
values given the species tree and the global parameter settings.  Because
recombinant edges from any species tree branch to the root branch are not
possible under \texttt{ClonalOrigin} model (see Figure S\ref{fig:clonalorigin}),
the column for the root as recipient was colored green. Similarly, all of the
cells where a recipient branch was older than the donor, and two branches did
not share any contemporary time were colored green. The diagonal cells were
colored gray, indicating no significant number of recombinant edges, because the
observed number of recombinant edges and the expected one were similar. As
recombinant edges starting from and ending in the same species tree branch would
not change the species tree, which would not change the likelihood much, the
cells should be colored gray.  Because SDD is a strictly veterinary pathogenic
bacteria, we did not expect that SDD received much gene transfer from either
SPY1 or SPY2: the log base 2 ratio from SPY1 to SDD was -2.20, and that from
SPY2 to SDD was -1.97. On the contrary, the recombing gene transfer from SDE1 to
SDD was 0.756, and that from SDE2 to SDD was 0.698. This relatively higher
recombining gene transfer from two SDE individuals to SDD compared with that
from two SPY individuals to SDD reflected that the SDD was more closely related
to SDE than to SPY in most of the genomes.

The genomic organization between SDD and SDE1/SDE2 were much similar (Figures
S\ref{fig:mauve}), and they formed a clade in the species tree
(S\ref{fig:tree5}). However, the gene transfer from SDD to either SDE1 or SDE2
was less frequent than the case of reverse direction; SDD served as a recipient
rather than a donor. It is not unclear why SDD received more genes from SDE1 or
SDE2 than did SDE1 or SDE2 from SDD.

The recombining gene transfer log base 2 ratios between SDE1 and SDE2 (1.17 and
1.20) were about twice as large as that from SDE1 and SDE2 to SPY1 and SPY2
(0.67, 0.66, 0.60, and 0.39).  The intensities of recombining gene transfer from
external branches of SDE to external branches to SPY was similar to those from
SDE1 or SDE2 to SDD (0.76 and 0.70).  Either SPY1 or SPY2, and SDD appeared to
be no different from the view point of SDE1 or SDE2 because of the similar
values of recombining gene transfer intensity from SDE1 or SDE2 to the other
three external branches.  It was expected that the recombining gene transfer
within a clade would be larger than that between clades.  Surprisingly, the
recombining gene transfer log base 2 ratios between SPY1 and SPY2 (1.06 and
1.07) was comparable with those from SPY1 and SPY2 to SDE1 and SDE2 (1.25, 1.25,
1.17, and 1.29).  Effectively, the recombining gene transfer from a clade SPY1
or SPY2 to a different clade SDE1 or SDE2 was no different from that within the
clade of SPY1 and SPY2. The two external branches of SDE received genes from
either SPY1 or SPY2 via recombining gene transfer as if SDE1 and SDE2 were in
the same clade as SPY1 and SPY2.  The bacterial recombination between SPY and
SDE might happen more often than that between SDE and SDD whereas the genome
configurations of SDE and SDD are more similar than each to SPY. This might be
consistent with the fact that SPY and SDE shared the same ecological niche.  The
gene transfer log base 2 ratio from SPY1 and SPY2 to SDE branch was larger than
any type of gene transfer: 1.88 and 1.94, respectively.  Furthermore, we found a
total of 58 genes for which a recombinant edge of certain types were sampled
with probability at least 0.9 by focusing on specific types of recombinant
edges.  (Table S\ref{tab:genes-transfer}).  Many of the genes in the list were
transferred from SPY to SDE rather than the reverse direction overall.  All
things considered, this clearly showed that the recombining gene transfer was
biased toward the net direction from SPY to SDE.

Figure \ref{fig:hgt-heatmap} shows the number of inferred non-recombining gene
transfer events between the different branches of the species tree.  SDE1 and
SDE2 shared the most events (26 and 34) compared to other sample pairs.  SDD
appeared to acquire many genes from the ancestral SDE branch. Transfer events
between the SPY and SDE clades were inferred in smaller numbers (<6) compared to
those within each clade (23--34). Among the clade-crossing transfers, we did not
see any significant enrichment for a specific direction, as opposed to the
result for recombining transfers from the model-based analysis (Figure
S\ref{fig:mowgli-recomb-heatmap}).

\subsection{More frequent recombining gene transfer than horizontal gene transfer}
Among the 2314 gene families
the largest family consisted of 59 genes, while
nearly half of gene families (1066) had exactly one gene per species
, or six genes total (Figure S\ref{fig:famsize}).
Figure \ref{fig:Gene-duplication-loss} shows the
inferred gene counts for each extant and ancestral species. Overall,
we inferred 304 duplications, 944 recombining gene transfers, and 294
non-recombining gene transfers, comprising 23.7\% of total transfer
events.  We found that duplications made up only a small fraction of
the events within the clade (10.2\%), while horizontal transfers and
losses were significantly more frequent (41.7\% and 48.0\%,
respectively). We found that duplications were most often
species-specific. Many gene losses likely represented horizontal
transfers that replaced the native gene sequence.  The species SDD and
SDE were the most expanded, both in terms of duplications and in-bound
transfers.

\subsection{Virulence genes via recombining gene transfer}

\texttt{Find cases of any virulence genes that were transferred via horizontally.} 
\texttt{Make a list of virulence genes and their categories with more interpretations.}

About 10\% of the genes comprised of virulence facters, about half of which were
found in the core genome of the five genome alignment. Using the UCSC genome
browser tracks we located virulence genes with notable posterior recombination
probability.  Stressful host conditions include heat shock, starvation, and iron
depletion.  Iron uptake of bacteria was known to be essential for survival of
the bacteria under hostile host condition. We found that parts of the gene,
putative calcium transporter (MgtA), in SDE1 had been transferred from SPY1
(SpyM3\_0440).  An extracellular protein, putative dipeptidase (SpyM3\_0465),
was found to be transferred from SPY to SDE. The gene as a whole was
transferred, which was a case for the finding that virulence genes tend to be
transferred as a whole unit \citep{Chan2009}. Dipeptidase may be important for
the survival of bacteria to scavenge dipeptides from the host.  An adhesins
gene, putative internalin A precursor (SpyM3\_1035), was found to be affected by
recombining gene transfer from SPY to SDE. Some of the middle parts of the gene
was affected.  A middle part of a chaperonin GroEL, putative heat shock protein
(SpyM3\_1765), was inferred to transfer from SPY to SDE1.  Increase of GroEL
expression was known to allow for the bacteria to survive the hostile host
environment of acid shock.  Superoxide dismutase was reported to be associated
virulence of \textit{Streptococcus agalactiae} \citep{Poyart2001}. This whole
gene including its 5' UTR, (SpyM3\_1071), was affected by recombining gene
transfer from SPY to SDE.  Pyruvate formate lyase was recently reported to have
virulence \citep{Yesilkaya2009}. A few parts of the gene (SpyM3\_1749) seemed to
be affected by recombining gene transfer.  One MLST locus, called DNA mismatch
repair protein MutS, appeared to be affected by recombining gene transfer.  A
putative spermidine/putrescine ABC transporter ATP-binding protein,
(SpyM3\_0764), was found to be transferred from SPY to SDE.  Under starvation
condition bacteria must use more effective energy metabolism.  A putative
glucose kinase, (SpyM3\_1180), might be transferred from SPY to SDE to adapt
SDE's energy metabolism to survive such harsh condition.

\subsection{Horizontal gene transfer is associated with virulence genes}
\texttt{Select genes with high horizontal gene transfers. Relate these to
categories of virulence genes. I should find many more virulence genes in Matt's
analysis.}

@@

@@

\subsection{Complete bacterial genomes}
Two genomes of \textit{S.\ pyogenes} (SPY), and two of \textit{S.\ dysgalactiae}
ssp.\textit{\ equisimilis} (SDE) were downloaded from NCBI with the following
accessions: NC\_004070 {Beres2002}, NC\_008024 \citep{Beres2006}, NC\_012891
\citep{Shimomura2011}, and CP002215 \citep{Suzuki2011}. We refer to these
samples as SPY1, SPY2, SDE1, and SDE2, respectively. One genome of 
\textit{S.\ dysgalactiae} ssp.\textit{\ dysgalactiae} was provided by 
\citet{Suzuki2011}
with accession identifiers of SddyATCC27957, referred to here as SDD. We also
downloaded the genome (NC\_012471) of \textit{S.\ equi} ssp.\textit{\ equi}
strain 4047 referred to as SEE \citep{Holden2009}.

@@

Two stages of MCMC with \texttt{ClonalOrigin}, subversion r19 that had been
downloaded from \url{http://clonalorigin.googlecode.com/svn/trunk}, were
employed independently for each of the 274 alignment blocks as described in
\citet{Didelot2010}.   In the first MCMC we obtained estimates of mutation rate,
recombination rate, and recombinant tract length.  Two independent chains were
executed with $10^6$ burn-in generations and $10^7$ sampling generations,
sub-sampled every $10^5$-th generations. One of the two chains was used to
obtain the block-wise mean estimates, and the other chain was used to validate
convergence of the first stage of MCMC.  The estimators of the three population
parameters were medians weighted over block lengths, which down-weighed less
reliable estimates obtained from shorter alignment blocks.

In the first stage of \texttt{ClonalOrigin} MCMC a few alignment blocks were not
successfully finished even after a month of computation.  Because those were not
expected to affect the global estimates of population parameters significantly,
we used only finished results of the first stage of \texttt{ClonalOrigin} to
estimate mutation rate, recombination rate, and recombinant tract length.  Two
independent runs of the first stage MCMC resulted in similar estimates of the
three population parameters (Table S\ref{tab:three}).

The second stage of
MCMC with \texttt{ClonalOrigin} was executed with the three population
parameters fixed to the estimated values in the first stage.  Two independent
chains of the second stage of \texttt{ClonalOrigin}'s MCMC were also executed
with a total generation being $1.1\times10^8$. The first $10^7$ generations of
each chain were discarded as burn-in. The remaining $10^8$ generations were
sub-sampled every $10^5$-th generation.

@@

@@
